PK object batch mode trial

### 2023-09-25 data-raw/batchMode_trial_gpm.Rmd
Updating vignette to prepare for publication, included a TOC for figure's produced.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
tidyverse_conflicts()
```

## Working with one DTXSID + Species combination

@@

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID8025337")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID9022360")

# A PFAS chemical?
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID3031860")
# All long lived chemicals
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c(long_lived_chem[3:5],
                                                              "DTXSID0020442"))
do_preprocess(my_pk_001p) %>% get_data() %>% filter(Time.Units != Time_trans.Units) %>%
  pull(Chemical) %>% unique() -> long_lived_chem

# Reasonable fits... multiple species
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID0020442",
                                                              "DTXSID2039336"))

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% (over1Ref %>%
                                                               pull(Chemical)),
                subjects.species_harmonized %in% "rat")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID0020442"),
                subjects.species_harmonized %in% "hamster",
                studies.administration_route_normalized %in% "iv")
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID1043881",
                                                              "DTXSID8020541",
                                                              "DTXSID6034849"),
                subjects.species_harmonized %in% "rat")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID3032179")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID4023917")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID7030066",
                                                              "DTXSID0020442",
                                                              "DTXSID2021315"),
                subjects.species_harmonized %in% "rat")

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID2039336",
                                                              "DTXSID7029904",
                                                              "DTXSID1021087",
                                                              "DTXSID2020682"))

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% c("DTXSID1034187",
                                                              "DTXSID2021315",
                                                              "DTXSID3032179",
                                                              "DTXSID4020533"))

@@

```{r single_fit}
# Pooled
my_pk <- pk(data = my_predata) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
# Add Reference for Joint Analysis

@@

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedConcentrations_2023-07-11.png"),
       height = 3.5,
       width = 5.5,
       units = "in")

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(DBI)
tidyverse_conflicts()
```

# TOC

1. Pulling Data from CvTdb  
  - Minimal `pk` object  
2. Running all possible fitting options  
  - Evaluating fitting options  
  - Head-to-head comparison of pooled vs joint models  
3. Running dose-normalized joint fits  
4. Analysis Plots  
  - CvTdb replicate variation (Figure 5)  
  - CvTdb replicate variation over ADME-normalized time (Figure 5B)  
  - CvTdb concentration and final time-point distributions (Supplementary Figure 1)  
  - invivoPKfit fit option selection (Figure 6)  
  - invivoPKfit model performance (Supplementary Figure 2)  
  - Model performance vs Data variability (Figure 7A)  
  - Goodness of fit plots (Figure 7B)  
  - Fits that may be improved (Supplementary Figure 3)  
5. Lombardo Analysis  
  - Chemicals included  
  - Comparison of derived TK stats  (Figure 8)
6. Benchmarking invivoPKfit
  - Parallelization speeds up the process of fitting (Supplementary Figure 4)

## Pulling data from CvTdb
This is largely taken from a `Querying CvTdb` vignette from Taylor Wall.

```{r query_functions}

postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
              user = Sys.getenv("CVT_USER"),
              password = Sys.getenv("CVT_PASS"),
              host = Sys.getenv("CVT_HOST"),
              dbname = Sys.getenv("CVT_DB_NAME"))
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) return(cat("\nMust provide a query to send"))
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN ", paste0("JOIN ", schema, "."), .))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

The following is the new SQL query, this will eventually be moved another vignette.

```{r SQL_queries}
# Sample query of CvTdb Documents SQL
# query <- "SELECT pmid, year, first_author, title FROM documents LIMIT 10"
# query <- "SELECT count(*) FROM studies"
# query <- "SELECT DISTINCT administration_route_normalized FROM studies"

# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  "a.id, a.time_original,  a.time_hr, a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",
  "b.id, b.analyte_name_original, b.analyte_dtxsid, b.analyte_casrn, ",
  "b.fk_analyzed_chemical_id, b.fk_test_chemical_id, ",
  "b.test_substance_dtxsid, b.analyte_name_secondary_original, ",
  "b.conc_medium_normalized, b.time_units_original, b.conc_units_original, ",
  "b.loq, b.loq_units, b.n_subjects_in_series, ",
  "c.id, c.administration_route_normalized, c.fk_dosed_chemical_id, c.dose_volume, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, c.dose_frequency, c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_original_units, ",
  "d.id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  "e.id, e.pmid, e.year, e.other_study_identifier, e.url, e.doi, e.extracted ",

  # Main Table
  "FROM conc_time_values a ",

  # Join with series table by series ID
  "LEFT JOIN series b ON b.id = a.fk_series_id ",

  # Join to studies table by study ID
  "LEFT JOIN studies c ON c.id = b.fk_study_id ",

  # Join to subjects table by subject ID
  "LEFT JOIN subjects d ON d.id = b.fk_subject_id ",

  # Join to documents table by extraction document ID
  "LEFT JOIN documents e ON e.id = c.fk_extraction_document_id ",

  # Filtering steps
  "WHERE b.conc_medium_normalized IN ('blood', 'plasma') AND ",
  "c.administration_route_normalized IN ('iv', 'oral')"
)

sql_query <- query_db(query = query,
         schema = "cvt")

names(sql_query)
# Need to change column names for this
{
  names(sql_query)[1] <- "conc_time_id"
  names(sql_query)[8] <- "series_id"
  names(sql_query)[22] <- "study_id"
  names(sql_query)[34] <- "subject_id"
  names(sql_query)[41] <- "document_id"
}

# Harmonize values for easier analysis
sql_query <- sql_query %>%
  mutate(
    species = tolower(species),
    sex = tolower(sex),
    loq = as.numeric(loq),
    n_subjects_in_series = as.numeric(n_subjects_in_series),
    dose_level_normalized = as.numeric(dose_level_normalized),
    dose_level_original = as.numeric(dose_level_original),
    dose_volume_units = str_extract(dose_volume,
                                    pattern = "(?<=[:space:]).+"),
    dose_volume = as.numeric(str_extract(dose_volume,
                                         pattern = "^(\\d|\\.)*")),
    dose_duration_units = str_extract(dose_duration,
                                      pattern = "(?<=[space]).+"),
    dose_duration = as.numeric(str_extract(dose_duration,
                                pattern = "^\\d"))
  )
sql_query <- sql_query %>%
  rename_with(~gsub(pattern = "fk_", replacement = "",
                         .x), .cols = contains("fk_"))

# Copying
cvt_df <- sql_query # Copy so I don't have to do query again to reset any errors

# Check other_study_identifier == S0916 so that it has dose_level_original_units == ug/mL
cvt_df %>% filter(other_study_identifier == "S0916") %>%
  pull(dose_level_original_units) %>% unique()
# There are still columns that are mg/kg... changing these to ug/mL
cvt_df <- cvt_df %>% mutate(
  dose_level_original_units = ifelse(
    other_study_identifier == "S0916",
    "ug/mL",
    dose_level_original_units
  )
)

# Need to convert the units to mg/kg
# In pulling_oral_iv_data_cvtdb.Rmd, it is noted that oral was 5mL/kg and iv 2mL/kg
# So I need to do ug/mL * mL/kg * 1mg/1000ug = mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0916",
      ifelse(administration_route_normalized %in% "oral",
             dose_level_original * 5/1000,
             dose_level_original * 2/1000),
      dose_level_normalized),
    dose_level_corrected_units = "mg/kg",
    .after = dose_level_original_units)

# For this study, it was in ng/kg not mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0624",
      dose_level_original * 1E-6,
      dose_level_normalized),
    .after = dose_level_original_units)

# The following measured C-14/H-3 but are indicators for the following Chemicals
# document_id = 65, DTXSID = DTXSID7031290, name = Isocyclemone E or beta-OTNE
# document_id = 61, DTXSID = DTXSID4022527, DTXSID3020209,
### name(s) = propylparaben, butylparaben, missing...methylparaben
# Propylparaben SERIES_ID = 110, Butylparaben SERIES_ID = 112
# For above, HPLC reviealed a single peak corresponding with p-hydroxybenzoic acid (PHBA, DTXSID3026647)
# document_id = 46, DTXSID = DTXSID2020139, name = benzo[a]pyrene
# document_id = 31, DTXSID = DTXSID2020139, name = benzo[a]pyrene

# However, now would be good time to eliminate any ones where they analyte != test chemical
# Some studies may have dosed with multiple chemicals... important to note
# cvt_df <- cvt_df %>% filter(analyzed_chemical_id == test_chemical_id)
# For now, I will change the chemical name and dtxsid
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 65, "DTXSID7031290",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 65, "beta-OTNE",
                                        analyte_name_original)) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "DTXSID4022527",
                                   series_id %in% 112 ~ "DTXSID3020209",
                                   .default = analyte_dtxsid
                                 ),
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "Propylparaben",
                                   series_id %in% 112 ~ "Butylparaben",
                                   .default = analyte_name_original
                                 ),
                                 analyte_name_original)
         ) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% c(46, 31),
                                 "DTXSID2020139",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% c(46, 31),
                                        "benzo[a]pyrene",
                                        analyte_name_original))

# For study S0976, Caffeine, L-ephedrine, and psuedoephedrine were dosed
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    other_study_identifier %in% "S0976",
    case_when(
      analyte_name_original %in% "Caffeine" ~ 4.00,
      analyte_name_original %in% "L-Ephedrine" ~ 4.44,
      analyte_name_original %in% "Pseudoephedrine" ~ 4.43
    ),
    loq),
    loq_units = ifelse(
      other_study_identifier %in% "S0976",
      "ng/mL",
      loq_units
    ))

# Some pentachloranisole studies
# S0361 seems to have some data that has not been extracted (also for mouse)
cvt_df <- cvt_df %>%
  mutate(loq = case_when(
    other_study_identifier %in% "S0361" ~ 0.05,
    other_study_identifier %in% "S0328" ~ 1, # This is inferred
    # Highest value under loq is 0.9
    # Lowest value above loq is 1.263, but there are differences in GC-EC and HPLC methods
    .default = loq
  ),
  loq_units = case_when(
    other_study_identifier %in% "S0361" ~ "ug/mL",
    other_study_identifier %in% "S0328" ~ "ug/mL",
    .default = loq_units
  ))

# Checking if entire groups of data are NA for conc and loq
# Really only eliminates Tamoxifen
# cvt_df %>%
#   # filter(analyte_dtxsid %in% "DTXSID1034187") %>%
#   group_by(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>%
#   filter(all(is.na(conc) & is.na(loq))) %>% 
#   distinct(analyte_dtxsid, document_id, analyte_name_original,
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>% View()

# Checking individual chemicals
# cvt_df %>%
#   filter(analyte_dtxsid %in% "DTXSID2020682") %>%
#   distinct(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected)

# PMID 6310077 is a tritium tracing of B[a]P and these were given as % of Dose
# The initial dose was 117 nmol, 2 mCi/kg
# Molecular Mass of benzo[a]pyrene (C20H12) is 252.316 g/mol
# The initial dose *I think* is in dose_level_normalized
cvt_df <- cvt_df %>%
  mutate(conc = ifelse(document_id %in% 38,
                       conc_original*dose_level_corrected,
                       conc))

# NTP-CEB C96019 just has a lot of data below LOQ
# For S0916, there is no > LOQ data for anything but Tamoxifen
# N-di-butyl phthalate is only good at the iv dose
# DTXSID2021315 is only good at 1E-4 Dose
# S0623 Doesn't have any plasma data
# There is a maybe control sample for hexachlorobenzene (Dose = 0.0 ) ?

# Filter by whether at each dose, there are no LOQ or Concentration value reported
# This is the threshold for Minimally Informative Data in our use case.
cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq))) %>%
  distinct(analyte_dtxsid, species,) %>% nrow() # Says there are 433 distinct

cvt_df <- cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq)))

# Give N-desmethyltamoxifen it's due DTXSID501315924
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "N-desmethyltamoxifen",
    "DTXSID501315924",
    analyte_dtxsid
  ))

cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "Methemoglobin",
    "DTXSID301056085",
    analyte_dtxsid
  ))

# Standardizing the names (example: glucuronide had two entries)
cvt_df <- cvt_df %>%
  mutate(analyte_name_original = tolower(analyte_name_original))

# Eliminate study that gave chemical each day, not single dose
# This may be used in another type of model, but right now we deal with single bolus
# Only include those that have NA or 1 for dose frequency
cvt_df <- cvt_df %>%
  filter(dose_frequency %in% c("1", NA))

# Normalize LOQ units
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    loq_units %in% c("ug/L", "ng/mL"), loq * 1E-3, loq)) # mg/L and ug/mL are same scale
# Normalize conc_sd and conc units manually
cvt_df <- cvt_df %>%
  mutate(conc_sd_normalized = case_when(
    conc_units_original %in% c("ug/L", "ng/mL", "ng-Eq/mL", "ng/g tissue") ~ conc_sd_original*1E-3,
    conc_units_original %in% c("g/dL") ~ conc_sd_original*1E-4,
    conc_units_original %in% c("pg/mL") ~ conc_sd_original*1E-6,
    .default = conc_sd_original
  ))
# Need to normalize the ng-eq/mL sample, conc is generally already normalized
cvt_df <- cvt_df %>%
  mutate(conc = case_when(
    conc_units_original %in% c("ng-eq/mL") ~ conc_original*1E-3,
    .default = conc
  ))

cvt_df <- cvt_df %>%
  dplyr::filter(!(analyte_dtxsid %in% c("DTXSID501315924",
                                      "DTXSID10169854",
                                      "DTXSID2021103") )) # These only have 1 or 2 values

# More notes:
# DTXSID6020438 & DTXSID0020868 rat is from water (top) and oil (bottom curve) gavage

```{r pk_obj_fromSQL}

### PK OBJECT ###----
# Minimum pk object, add options later
minimal_pk <- pk(data = cvt_df,
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
```

## Running all possible fitting options

# Pooled
my_pk_000p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_100p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_110p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_111p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_010p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_011p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_001p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_101p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

```{r joint_fitting_choices}
# Joint
my_pk_000j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_100j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_110j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_111j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_010j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_011j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_001j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_101j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
```

### Evaluating fitting options

  cli::cli_progress_step("Winning Model...")
  # Wide winning model
  suppressMessages({
    winning_model <- get_winning_model(obj = obj) 
    wide_winning_model <- winning_model %>%
      group_by(method, model) %>%
      count() %>%
      pivot_wider(names_from = model, values_from = n) %>%
      mutate(Options = pk_name)
  })

  my_df <- left_join(fold_errors_all, r2_df) %>%
    inner_join(myAIC) %>%
    inner_join(myRMSE_all) %>%
    mutate(Options = pk_name, .before = Chemical)

  full_list <- list(wide_winning_model, RMSE_eval, my_df)
  })

### Evaluations-----
my_pk_000p <- do_fit(my_pk_000p, n_cores = 14) 
my_pk_000p <- my_pk_000p %>% evaluate_choices() # Done
my_pk_100p <- do_fit(my_pk_100p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_110p <- do_fit(my_pk_110p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_111p <- do_fit(my_pk_111p, n_cores = 8) %>% evaluate_choices() # Done
rm(my_pk_100p,my_pk_110p,my_pk_111p)
gc()
my_pk_010p <- do_fit(my_pk_010p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_011p <- do_fit(my_pk_011p, n_cores = 8) %>% evaluate_choices() # Done 
my_pk_001p <- do_fit(my_pk_001p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_101p <- do_fit(my_pk_101p, n_cores = 8) %>% evaluate_choices() # Done

my_pk_000j <- do_fit(my_pk_000j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_100j <- do_fit(my_pk_100j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_110j <- do_fit(my_pk_110j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_111j <- do_fit(my_pk_111j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
rm(my_pk_000j, my_pk_100j, my_pk_110j, my_pk_111j)
gc()
my_pk_010j <- do_fit(my_pk_010j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_011j <- do_fit(my_pk_011j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_001j <- do_fit(my_pk_001j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_101j <- do_fit(my_pk_101j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done

# Run after each evaluation
# initialization (only run once)
{
winning_model_final <- NULL
rmse_summary <- NULL
detail_stats <- NULL
}

# Run all at once
{
current_pk <- my_pk_101j # Define current evaluated pk object here
winning_model_final <- bind_rows(winning_model_final, current_pk[[1]])
rmse_summary <- bind_rows(rmse_summary, current_pk[[2]])
detail_stats <- bind_rows(detail_stats, current_pk[[3]])
rm(current_pk)
gc()
}

## Head-to-head comparison of pooled vs joint models

Now I need to go back and re-analyze the data only for 100, 101, and 000
Pooled and Joint, L-BFGS-B only. I need to filter the cvt_df to only include Chem + Species combinations
with multiple references. Then I will join them with the excel sheet of already collected data.

```{r load_multirefs}
# Reference Multiples
multiRef_ChemSpec <- cvt_df %>%
  ungroup() %>%
  distinct(analyte_dtxsid, species, document_id) %>%
  group_by(analyte_dtxsid, species) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  dplyr::select(!n)
multiRef_ChemSpec %>% nrow() # Surprisingly only 13 Chem + Species have multiple references

```{r Plotting_Summary}
finalist_options <- c("000 Pooled", "000 Joint",
                      "100 Pooled", "100 Joint",
                      "101 Pooled", "101 Joint")

readxl::excel_sheets(paste(Sys.getenv("OD_DIR"),
                           "DataSummary_9_21_2023.xlsx",
                           sep = "/")) # Predictions Evaluation

preval <- readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Predictions Evaluation") %>%
  left_join(multiRef_ChemSpec,
            by = join_by(Chemical == analyte_dtxsid,
                         Species == species)) %>%
  filter(method %in% "L-BFGS-B", 
         Options %in% finalist_options) %>%
  distinct()

preval %>%
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  ggplot(aes(x = avg_FoldErr,
             color = Options)) +
  geom_freqpoly(bins = 10) +
  facet_grid(cols = vars(Error_Model))

preval %>% 
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  group_by(Options, Error_Model) %>%
  summarize(totalFoldError = sum(avg_FoldErr))

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Fit_Counts") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Cmax and AUC_inf Evaluation") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

## Running dose-normalized joint fits

Dose-normalization and a hierarchical model were the options that produced the least
amount of error in predictions and derived TK stats.

```{r Standard_Settings}
# Full option setting
my_pk <- pk(data = cvt_df,
            mapping = ggplot2::aes(
              Chemical = analyte_dtxsid,
              Chemical_Name = analyte_name_original,
              DTXSID = analyte_dtxsid,
              CASRN = analyte_casrn,
              Species = species,
              Reference = document_id,
              Media = conc_medium_normalized,
              Route = administration_route_normalized,
              Dose = dose_level_normalized,
              Dose.Units = "mg/kg",
              Subject_ID = subject_id,
              Series_ID = series_id,
              Study_ID = study_id,
              ConcTime_ID = conc_time_id,
              N_Subjects = n_subjects_in_series,
              Weight = weight_kg,
              Weight.Units = "kg",
              Time = time_hr,
              Time.Units = "hours",
              Value = conc,
              Value.Units = "mg/L",
              Value_SD = conc_sd_normalized,
              LOQ = loq
            )) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
my_pk <- do_preprocess(my_pk)
my_data <- get_data(my_pk)

An aside: working with one DTXSID + Species combination

@@

@@

```{r single_fit}
my_pk_original <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
# Add Reference for Joint Analysis

my_pk_original <- do_preprocess(my_pk_original)

@@

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedConcentrations_2023-09-21.png"),
       height = 3.5,
       width = 5.5,
       units = "in")
